The classical mutual information in mean-field spin glass models

### 2015-07-15 sk_model_v18.tex
Small typo corrections

All these issues can be thoroughly addressed in the Sherrington-Kirkpatrick 
(S-K) model~\cite{sherrington-1978,sherrington-1978-prl}, which is exactly 
solvable. The S-K model is a \emph{classical} Ising model on the fully-connected 
graph of $N$ sites, with quenched random interactions. Its hamiltonian reads  
%
\begin{equation}
{\mathcal H}=-\sum\limits_{1\le i<j\le N}J_{ij}S_i S_j-
h\sum\limits_{1\le i\le N}S_i.
\label{SK-intro}
\end{equation}
%
Here $S_i=\pm 1$ are classical Ising spins, $h$ is an external magnetic field, and 
$J_{ij}$ are uncorrelated (from site to site) random variables. The S-K model hosts 
a low-temperature glassy phase, which is separated from a high-temperature paramagnetic 
one by a second order phase transition. Despite its mean-field nature, the solution 
of the S-K model has been a mathematical challenge. Although it was proposed as an 
ansatz by Parisi~\cite{parisi-1980} more than thirty years ago, its rigorous proof 
was obtained only recently~\cite{talagrand-2006}. Moreover, the solution exhibits several 
intricate features, such as lack of self-averaging~\cite{pastur-1991}, ultrametricity~\cite{
mezard-1984,rammal-1986}, and replica symmetry breaking~\cite{parisi-book,castellani-2005}. 
The last refers to the breaking of the permutation symmetry among the fictitious replicas 
of the model, which are introduced to perform disorder averages (via the so-called 
\emph{replica trick}~\cite{cardy-book}. Finally, although the 
applicability of the S-K model to describe realistic spin glasses is still highly 
debated~\cite{yucesoy-2012,billoire-2012,yucesoy-2013}, there are recent proposals on 
how to realize it in cold-atomic gases~\cite{morrison-2008,rotondo-2015}, or in laser 
systems~\cite{ghofraniha-2015}.

In the last decade entanglement-related quantities emerged as valuable tools to 
understand the physics of complex systems~\cite{amico-2008,eisert-2009,calabrese-2009,
cc-rev}, both classical and quantum. For instance, at a conformally invariant critical 
point entanglement measures contain universal information about the underlying 
conformal field theory (CFT), such as the central charge~\cite{holzhey-1994,vidal-2003,
calabrese-2004,calabrese-2012}. For classical spin models a lot of attention has been 
focused on the \emph{classical} R\'enyi entropy~\cite{jaconis-2013,stephan-2014}. Given 
a bipartition of the system into two complementary subregions $A$ and $B$, the classical 
R\'enyi entropy $S_n(A)$ (with $n\in\mathbb{N}$) is defined as 
%
\begin{equation}
S_n(A)\equiv \frac{1}{1-n}\log\Big(\sum\limits_{i_A\in{\mathcal C}_A} p^n_{i_A}
\Big)
\label{renyi-intro}
\end{equation}
%
Here ${\mathcal C}_A$ denotes the set of all the possible spin configurations in part 
$A$, whereas $p_{i_A}$ is the probability of the configuration $i_A$. 
Alternatively, $S_n(A)$ can be obtained from the partition function of the model on 
an \emph{ad hoc} defined ``booklet'' geometry (see section~\ref{booklet} for its 
definition). This consists of $n$ independent and identical copies (the 
booklet ``sheets'') of the model. These \emph{physical} copies are different 
from the \emph{fictitious} replicas used to perform the disorder average. Each sheet is 
divided into two parts $A$ and $B$, containing $N_A$ and $N_B$ spins, respectively. 
The spins in part $A$ of different sheets are constrained to be equal. It is 
convenient to introduce the booklet aspect ratio $0\le\omega\le1$ as  
%
\begin{equation}
\label{a-ratio}
\omega\equiv \frac{N_A}{N}.
\end{equation}
%
Notice that Eq.~\eqref{renyi-intro} can also be used for quantum systems by replacing 
the sum over the proababilities of each state in region $A$ with the trace over the reduced 
density matrix for region $A$. For $n=1$ 
Eq.~\eqref{renyi-intro} defines the subsystem Shannon entropy~\cite{alcaraz-2013,
stephan-2014-a}. From $S_n(A)$, one defines the classical mutual information 
${\mathcal I}_n(A,B)$ as 
%
\begin{equation}
\label{MI}
{\mathcal I}_n(A,B)\equiv S_n(A)+S_n(B)-S_n(A\cup B). 
\end{equation}
%
For local models ${\mathcal I}_n$ obeys the area law~\cite{wolf-2008} ${\mathcal I}_n(A)
\propto\ell$, with $\ell$ the length of the boundary between $A$ and $B$. Remarkably, 
for different $\ell$, the ratio ${\mathcal I}_n/\ell$ has been shown to exhibit a crossing at a second 
order phase transition~\cite{jaconis-2013}, implying that it can be used as a diagnostic 
tool for critical behaviors. For conformally invariant critical models more universal 
information can be extracted from the area-law corrections of ${\mathcal I}_n$~\cite{stephan-2014}.

>>>>

All these issues can be thoroughly addressed in the Sherrington-Kirkpatrick 
(S-K) model~\cite{sherrington-1978,sherrington-1978-prl}, which is exactly 
solvable. The S-K model is a \emph{classical} Ising model on the fully-connected 
graph of $N$ sites, with quenched random interactions. Its hamiltonian reads  
%
\begin{equation}
{\mathcal H}=-\sum\limits_{1\le i<j\le N}J_{ij}S_i S_j-
h\sum\limits_{1\le i\le N}S_i.
\label{SK-intro}
\end{equation}
%
Here $S_i=\pm 1$ are classical Ising spins, $h$ is an external magnetic field, and 
$J_{ij}$ are uncorrelated (from site to site) random variables. The S-K model hosts 
a low-temperature glassy phase, which is separated from a high-temperature paramagnetic 
one by a second order phase transition. Despite its mean-field nature, the solution 
of the S-K model has been a mathematical challenge. Although it was proposed as an 
ansatz by Parisi~\cite{parisi-1980} more than thirty years ago, its rigorous proof 
was obtained only recently~\cite{talagrand-2006}. Moreover, the solution exhibits several 
intricate features, such as lack of self-averaging~\cite{pastur-1991}, ultrametricity~\cite{
mezard-1984,rammal-1986}, and replica symmetry breaking~\cite{parisi-book,castellani-2005}. 
The last refers to the breaking of the permutation symmetry among the fictitious replicas 
of the model, which are introduced to perform disorder averages (via the so-called 
\emph{replica trick}~\cite{cardy-book}). Finally, although the 
applicability of the S-K model to describe realistic spin glasses is still highly 
debated~\cite{yucesoy-2012,billoire-2012,yucesoy-2013}, there are recent proposals on 
how to realize it in cold-atomic gases~\cite{morrison-2008,rotondo-2015}, or in laser 
systems~\cite{ghofraniha-2015}.

In the last decade entanglement-related quantities emerged as valuable tools to 
understand the physics of complex systems~\cite{amico-2008,eisert-2009,calabrese-2009,
cc-rev}, both classical and quantum. For instance, at a conformally invariant critical 
point entanglement measures contain universal information about the underlying 
conformal field theory (CFT), such as the central charge~\cite{holzhey-1994,vidal-2003,
calabrese-2004,calabrese-2012}. For classical spin models a lot of attention has been 
focused on the \emph{classical} R\'enyi entropy~\cite{jaconis-2013,stephan-2014}. Given 
a bipartition of the system into two complementary subregions $A$ and $B$, the classical 
R\'enyi entropy $S_n(A)$ (with $n\in\mathbb{N}$) is defined as 
%
\begin{equation}
S_n(A)\equiv \frac{1}{1-n}\log\Big(\sum\limits_{i_A\in{\mathcal C}_A} p^n_{i_A}
\Big)
\label{renyi-intro}
\end{equation}
%
Here ${\mathcal C}_A$ denotes the set of all the possible spin configurations in part 
$A$, whereas $p_{i_A}$ is the probability of the configuration $i_A$. 
Alternatively, $S_n(A)$ can be obtained from the partition function of the model on 
an \emph{ad hoc} defined ``booklet'' geometry (see section~\ref{booklet} for its 
definition). This consists of $n$ independent and identical copies (the 
booklet ``sheets'') of the model. These \emph{physical} copies are different 
from the \emph{fictitious} replicas used to perform the disorder average. Each sheet is 
divided into two parts $A$ and $B$, containing $N_A$ and $N_B$ spins, respectively. 
The spins in part $A$ of different sheets are constrained to be equal. It is 
convenient to introduce the booklet aspect ratio $0\le\omega\le1$ as  
%
\begin{equation}
\label{a-ratio}
\omega\equiv \frac{N_A}{N}.
\end{equation}
%
Notice that Eq.~\eqref{renyi-intro} can also be used for quantum systems by replacing 
the sum over the probabilities of each state in region $A$ with the trace over the reduced 
density matrix for region $A$. For $n=1$ 
Eq.~\eqref{renyi-intro} defines the subsystem Shannon entropy~\cite{alcaraz-2013,
stephan-2014-a}. From $S_n(A)$, one defines the classical mutual information 
${\mathcal I}_n(A,B)$ as 
%
\begin{equation}
\label{MI}
{\mathcal I}_n(A,B)\equiv S_n(A)+S_n(B)-S_n(A\cup B). 
\end{equation}
%
For local models ${\mathcal I}_n$ obeys the area law~\cite{wolf-2008} ${\mathcal I}_n(A)
\propto\ell$, with $\ell$ the length of the boundary between $A$ and $B$. Remarkably, 
for different $\ell$, the ratio ${\mathcal I}_n/\ell$ has been shown to exhibit a crossing at a second 
order phase transition~\cite{jaconis-2013}, implying that it can be used as a diagnostic 
tool for critical behaviors. For conformally invariant critical models more universal 
information can be extracted from the area-law corrections of ${\mathcal I}_n$~\cite{stephan-2014}.
